---
title: "Water"
output: html_notebook
---


Examples from [Statistical Rethinking](https://xcelab.net/rm/statistical-rethinking/). Markdown [cheatsheet](https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)

To render html in GitHub prefix `https://htmlpreview.github.io/?` to URL.

### Data Story

* Globe with proportion $p$ of water $W$

* Throw & catch globe and put finger at random place has probability     $p$ of water $W$ or probability $(1-p)$ of land $L$

* Tosses are independent

### Grid approximation of posterior

Likelihood is binomial distributed.
```{r}
plot_posterior_grid <- function(prior) {
  likelihood <- dbinom(6, size=9, prob=p_grid)
  unstd.posterior <- likelihood * prior
  posterior <- unstd.posterior / sum(unstd.posterior)
  plot(p_grid,posterior,type='b', xlab='probability of water', ylab='posterior probability')
  mtext("20 points")
}

plot_posterior_grid( rep(1,20) )
```

I know world has more than 50% qwater:
```{r}
prior <- ifelse( p_grid < 0.5, 0, 1 )
plot_posterior_grid( prior )
```

If I believed 50% is most likely:
```{r}
prior <- exp( -5*abs( p_grid - 0.5 ) )
plot_posterior_grid( prior )
```
